{\helveticaitalic{Supplementary Material}}

### 2022-02-04 tex/sections/introduction.tex
Split latex document into subfiles

\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{soul}
\usepackage{physics}
\usepackage{bm}

\usepackage{url,hyperref,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}
\usepackage{cleveref}
% \usepackage[font={bold,it}]{caption}

% Define better ref abbreviations
\usepackage[displaymath]{lineno}

\renewcommand{\d}{\mathrm{d}}
\renewcommand{\k}{\mathrm{K}}
\newcommand{\ca}{\mathrm{Ca}}
\newcommand{\na}{\mathrm{Na}}
\newcommand{\leak}{\mathrm{L}}
\newcommand{\dstar}{d^\star}
\newcommand{\gstar}{g^\star}
\newcommand{\gbar}{\bar g}
\newcommand{\delt}{\Delta t}
\newcommand{\taus}{\tau_s}
\newcommand{\dn}{\delta_n}
\newcommand{\taud}{\tau_d}

\linenumbers

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Olenik {et~al.}} %use et al only if is more than 1 author
\def\Authors{Mark Olenik\,$^{1,*}$, and Conor Houghton\,$^{2}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$School of Biological Sciences, Faculty of Life Sciences, University of Bristol, Bristol, United Kingdom \\
$^{2}$School of Computer Science, Electrical and Electronic Engineering, and Engineering Mathematics, Faculty of Engineering, University of Bristol, Bristol, United Kingdom}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Mark Olenik}

\def\corrEmail{m.olenik@bristol.ac.uk}

\begin{document}

\onecolumn
\firstpage{1}

\title[Scalar Poincaré map for bursting]{A scalar Poincaré map for anti-phase bursting in coupled inhibitory neurons with synaptic depression.}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}
\section{}
Short-term synaptic plasticity is found in many areas of the central
nervous system.  In the inhibitory half-centre central pattern
generators involved in locomotion, synaptic depression is believed to
act as a burst termination mechanism, allowing networks to generate
anti-phase bursting patterns of varying periods.  To better understand
burst generation in these central patter generators, we study a
minimal network of two neurons coupled through depressing synapses.
Depending on the strength of the synaptic conductance between the two
neurons, this network can produce symmetric $n-n$ anti-phase bursts,
where neurons fire $n$ spikes in alternation, with the period of such
solutions increasing with the strength of the synaptic conductance.
Relying on the timescale disparity in the model, we reduce the
eight-dimensional network equations to a fully-explicit scalar
Poincaré burst map.  This map tracks the state of synaptic depression
from one burst to the next and captures the complex bursting dynamics
of the network.  Fixed points of this map are associated with stable
burst solutions of the full network model, and are created through
fold bifurcations of maps.  We derive conditions that describe
period-increment bifurcations between stable $n-n$ and $(n+1)-(n+1)$
bursts, producing a full bifurcation diagram of the burst cycle
period.  Predictions of the Poincaré map fit excellently with
numerical simulations of the full network model and allow the study of
parameter sensitivity for rhythm generation.

\tiny
 \keyFont{ \section{Keywords:} Synaptic depression, Poincaré map, Dynamical system, Neuronal bursting, Central pattern generator}

\end{abstract}

\section{Introduction}
% DEPRESSION
Short-term synaptic plasticity may have a role in burst activity in
central pattern generators (CPGs).  Short-term synaptic depression is
commonly found in neuronal networks involved in the generation of
rhythmic movements, such as in the pyloric CPG of the spiny
lobster~\citep{manor1997, rabbah2007}, or in the lumbosacral cord of
the chick embryo~\citep{donovan1998}.  Synaptic depression modulates
the strength of synapses in response to changes to the presynaptic
firing frequency.  At a high neuronal firing frequency, depression
weakens the strength of synapses and therefore reduces the magnitude
of the postsynaptic response.  At low firing frequency, it allows
sufficient time for the synapse to recover from depression between
spikes, leading to a stronger postsynaptic response.  In reciprocal
networks, synaptic depression has been shown to act as a ``switch'',
giving rise to a wide range of network dynamics such as synchronous
and multi-stable rhythms, as well as fine tuning the frequency of
network oscillations~\citep{nadim2000, nadim1999, bose2011}.

@@

\section{Materials and Methods}
We consider a pair of identical Morris-Lecar neurons~\citep{morris1981}, with parameters adapted from~\cite{bose2011}.
The Morris-Lecar model is a set of two first-order differential equations that describe the membrane dynamics of a spiking neuron.
The depolarisation is modelled by an instantaneous calcium current, and the hyperpolarisation by a slow potassium current and a leak current.
The membrane potential $v_{i}$ and potassium activation $w_{i}$ of neuron $i$ ($i, j=1,2$) is described by:
\begin{align}
    ~\label{eq:cell-modelA}
     \dot v_{i}&= f(v_{i}, w_{i}) -\gbar s_j(v_i-v_{s}),\\
    ~\label{eq:cell-modelB}
     \dot w_{i} &=h(v_i,w_i).
   \end{align}
Here $v_{s}$ is the inhibitory reversal potential, and $\gbar$ and $s_{j}$ are the maximal synaptic conductance and the synaptic gating, respectively, constituting the total inhibitory conductance $\gbar s_{j}$ from neuron $j$ to neuron $i$.
Function $f(v_{i}, w_{i})$ describes the membrane currents of a single cell:
\begin{equation}
 ~\label{eq:memcur}
  f(v_{i}, w_{i}) = -g_{\ca}m_{\infty}(v_{i})(v_{i}-v_{\ca}) - g_{\k}w_{i}(v_{i}-v_{\k})-g_{\leak}(v_{i}-v_{\leak}) + I.
\end{equation}
The currents include a constant current $I$, and three ionic currents:
an instantaneous calcium current, a potassium current, and a leak current, with respective reversal potentials $v_{\ca}$, $v_{\k}$, and $v_{\leak}$, as well as maximum conductances $g_{\ca}$, $g_{\k}$, and $g_{\leak}$.
The function $h(v_{i}, w_{i})$ models the kinetics of the potassium gating variable $w_{i}$, and is given by
\begin{equation}
  h(v_{i}, w_{i})=\frac{w_{\infty}(v_{i})-w_{i}}{\tau_{w}}~\label{eq:h}.
\end{equation}
The steady-state activation functions $m_{\infty}$ and $w_{\infty}$ as well as the default model parameters are described in the \textbf{Supplementary Material S1}.

The dynamics of the synaptic interactions between the neurons are governed by a synaptic gating variable $s_{i}$ and a depression variable $d_{i}$:
\begin{linenomath}
  \begin{align}
    \dot s_{i} &= -s_{i}/\tau_{s}~\label{eq:dot-s},\\
    \dot d_{i} &= (1-d_{i})/\tau_{d}~\label{eq:dot-d}.
  \end{align}
\end{linenomath}
Variable $d_{i}$ describes a firing rate dependent depletion mechanism that governs the amount of depression acting on the synapse.
Our model is agnostic with respect to the exact mechanism of this depletion, be it pre- or post-synaptic.
In between spikes $s_{i}$ decays with time constant $\tau_{s}$, while $d_{i}$ recovers with time constant $\tau_{d}$.
Because synaptic depression occurs on a much slower timescale than synaptic inhibition, we assume $\tau_{d}\gg\tau_{s}$.
At spike time when the voltage $v_{i}$ crosses a synaptic threshold $v_{\theta}$, the synaptic variable $s_{i}$ is reset to the current value of $d_{i}$, and $d_{i}$ is scaled down by the depression strength parameter $\lambda$, where $0<\lambda<1$.
This results in the following discontinuous reset rule on spike time:
\begin{linenomath}
  \begin{align}
    s_{i} &\leftarrow d_{i}~\label{eq:s-reset},\\
    d_{i} &\leftarrow \lambda d_{i}~\label{eq:d-reset}.
  \end{align}
\end{linenomath}
The equations for the depression model were adapted from the~\citet{bose2001} model.
These equations are a mathematically tractable simplification of the established phenomenological depression model previously described by~\citet{tsodyks1997}.
The original equations in~\cite{bose2001} include an exponential decay of $d_{i}$ throughout the active phase of the action potential, that is $\dot d_{i}=-d_{i}/\tau_{\beta}$ for $v_{i}>v_{\theta}$.
To simplify the mathematical analysis we have reduced these dynamics to a single-reset event at spike time according to \cref{eq:d-reset}.
This simplification is natural given an approximately constant duration of the active phase of an action potential, and it will become evident from our results that this simplification does not qualitatively change the dynamics of the network compared to the original equations, where depression dynamics are described by an exponential decay.

When the cells are uncoupled ($\gbar=0$), the membrane dynamics are determined by the cubic $v$-nullcline $v_{\infty}(v_i)$ and the sigmoid $w$-nullcline $w_{\infty}(v_{i})$, satisfying $\dot v_{i}=0$ and $\dot w_{i}=0$, respectively.
The two curves intersect along the middle branch of $v_{\infty}$, creating an unstable fixed point $p_{f}=(v_{f},w_{f})$ with a surrounding stable limit cycle of period $T$ (\cref{fig:nullclines}A).
Trajectories along that limit cycle have the familiar shape of the action potential (\cref{fig:nullclines}B).
The trajectory of an action potential can be dissected into four phases: (1) a silent phase, (2) a jump up, (3) an active phase, and (4) a jump down~\citep[see e.g.][]{ermentrout2010}.
During the silent phase the trajectory evolves along the left branch ($v_{i}<v_{\theta}$) of the cubic $v$-nullcline.
Once the trajectory reaches the local minimum of $v_{\infty}$, it ``jumps up'' to the right branch ($v_{i}>v_{\theta}$), crossing the firing threshold $v_{\theta}$.
During the active phase the trajectory then evolves along the right branch of the cubic until it arrives at the local maximum, where it ``jumps down'' to the left branch commencing a new cycle.

The two-cell network model is numerically integrated using an adaptive step-size integrator for stiff differential equations implemented with XPPAUT~\citep{ermentrout2002} and controlled through the Python packages SciPy~\citep{scipy2020} and PyXPP~\citep{pyxpp}.
The following mathematical analysis is performed on the equations of a single cell.
Unless required for clarity, we will therefore omit the subscripts $i,j$ from here on.

\section{Results}
\subsection{Anti-phase burst solutions}
Short-term synaptic depression of inhibition in a half-centre oscillator acts as a \emph{burst termination} mechanism~\citep{brown1911} and is known to produce $n-n$ anti-phase burst solutions of varying period.
Such $n-n$ solutions consist of cells firing bursts of $n$ spikes in alternation.
\Cref{fig:depression-traces}A shows the timecourse of a typical $4-4$ burst.
While one cell is firing a burst it provides an inhibitory conductance to the other cell,  preventing it from firing.
Therefore, at any given moment one cell is spiking while the other is inhibited.
Consistent with~\citet{bose2011} we will refer to the currently firing cell as ``free'' and we will call the inhibited cell ``quiet''.
Additionally, we will distinguish between two phases of a $n-n$ solution:
We will refer to the burst duration of a cell as the ``free phase'', which is the time between the first spike and the last spike in a burst. And we will call the remaining duration of a cycle, when a cell is not spiking, the ``quiet phase''.

% TODO: Add mention of gstar. Once critical conductance is crossed, cell is released.
With each action potential of the free cell, short-term depression leads to a step-wise decrease of $d$, and consequently of $s$ (\cref{fig:depression-traces}B).
If $d$ depresses faster at spike time than it can recover in the inter-spike-intervals ($ISI$s), the total synaptic conductance $\gbar s$ will eventually become sufficiently small to allow for the quiet cell to be released and start firing, thus inhibiting the previously free cell.
While a cell is quiet its depression variable can recover.
Once the quiet cell becomes free again its synaptic inhibition will be sufficient to terminate the burst of the previously free cell and commence a new cycle.
As previously demonstrated by~\citet{bose2011}, in a two-cell reciprocally inhibitory network with synaptic depression the coupling strength $\gbar$ determines the type of $n-n$ solution.
Increasing $\gbar$ produces higher $n-n$ burst solutions with more spikes per burst and a longer cycle period.
\Cref{fig:burst-sols} shows numerically stable $n-n$ solutions for varying values of $\gbar$.
For small values of $\gbar$ the network produces anti-phase spiking $1-1$ solutions.
As $\gbar$ is increased the network generates solutions of increasing $n$, that is $2-2$, $3-3$, and $4-4$.
When $\gbar$ is sufficiently large (bottom of \cref{fig:burst-sols}), one of the cells continuously spikes at its uncoupled period $T$ while the other cell remains fully suppressed.
Depending on the initial conditions either of the two cells can become the suppressed cell, which is why the suppressed solution is numerically bistable.

Branches of numerically stable $n-n$ solutions and their associated limit cycle period for varying values of $\gbar$ are depicted in \cref{fig:bif-diagram}A (see  \textbf{Supplementary Material S2} for algorithm description).
Not only do higher $n-n$ solutions branches require stronger coupling $\gbar$, but also within $n-n$ branches the period increases with $\gbar$.
In line with~\citet{bose2011} we find small overlaps between solution branches indicating numerical bistability, for example such as between the $2-2$ and $3-3$ solution branches.
Branches of higher $n-n$ burst solutions occur on increasingly smaller intervals of $\gbar$, for instance is the $\gbar$ interval of the $5-5$ branch shorter than that of the $4-4$ branch and so on.
The interval between the $5-5$ branch and the suppressed solution (region between dotted lines in \cref{fig:bif-diagram}A) not only contains even higher numerically stable $n-n$ solutions, such as $11-11$ bursts, but also other non-symmetric $n-m$ solutions as well irregular, non-periodic solutions. However, the analysis in the following sections will only be concerned with the numerically stable and symmetric $n-n$ solutions.

\subsection{Mathematical analysis of two-cell network}
~\label{sec:assumptions}
% Summary of assumptions
The goal of the following mathematical analysis is to reduce the complexity of the eight-dimensional system to some easily tractable quantity.
As we will see later this quantity is the value of the depression variable $d$ of either of the two cells.
We will construct the solution of $d$ in a piecewise manner from one spike to the next, first during the free phase, and then during the quiet phase.
This construction will require two assumptions about the membrane and synaptic dynamics.
The first assumption states that during a burst the free cell fires at its uncoupled period $T$, which simplifies the construction of the solution of $d$.
The second assumption states that once the inhibitory conductance acting on the quiet cell drops below a critical threshold, the cell is immediately released and fires.
The second assumption is necessary to predict the release time of the quiet cell, which allows us to model the recovery of $d$ during the quiet phase.
In other words, the second assumption requires that the release of the quiet cell from inhibition depends only on the timecourse of the inhibition, and not on the membrane dynamics of the quiet cell.
Both assumptions can be observed in coupled relaxation-oscillator types of neurons such as the Morris-Lecar model we use, and will be numerically verified below.
Both assumptions were first explored in~\cite{bose2011} to derive algebraic conditions that guarantee the periodicity of the depression variable for different $n-n$ solutions.
However here we will use these assumptions to construct a Poincaré map of $d$, which will provide a geometric intuition for the dynamics of the full two-cell network and its dependence on model parameters.

% Assumption 1: Free phase
Our first assumption about the model states that the free cell fires at its uncoupled period $T$, that is, during the free phase of a burst we have $ISI=T$.
Solution profiles in \cref{fig:burst-sols} suggest that the $ISI$s are indeed approximately constant.
We can further numerically confirm this observation by capturing the $ISI$s of the stable solutions from the bifurcation diagram in \cref{fig:bif-diagram}A.
In addition to $ISI$s, \Cref{fig:bif-diagram}B also shows inter-\textit{burst} intervals ($IBI$s), which correspond to the time interval between the last spike of the burst and the full cycle period, and which lie in the quiet phase of the burst.
$IBI$s lie on multiple branches, each branch associated with a stable $n-n$ solution, and are monotonically increasing with $\gbar$.
In contrast, $ISI$s are calculated from the spikes within the free phase and do not vary significantly with $\gbar$, but are approximately $ISI\approx T$, affirming our first assumption.
Assuming $ISI=T$ allows us to ignore the non-linear membrane dynamics during the free phase, and to construct the evolution of the synaptic variables iteratively from spike to spike.
Assuming $ISI=T$ seems reasonable given that inhibition acting on the quiet cell decays exponentially to zero on a much shorter timescale than the duration of the $ISI$, and therefore, once the quiet cell is released its trajectory quickly approaches the spiking limit cycle.

% Assumption 2: Release
Our second assumption states that the quiet cell is released and spikes as soon as inhibition from the free cell drops below a constant threshold.
We will now define such a ``release condition'' by exploiting the discrepancy in timescales between the fast membrane dynamics, and the slower synaptic dynamics.
Let us first consider the dynamics of a single Morris-Lecar neuron.
We fix the synaptic variable that acts on the cell by setting $s=1$, which also makes the applied synaptic conductance $\gbar s$ constant.
Recall from \cref{fig:nullclines}A that in case of a single uncoupled cell ($\gbar s=0$), the $v$- and $w$-nullclines intersect at some unstable fixed point $p_{f}=(v_{f},w_{f})$, while trajectories revolve around a stable spiking limit cycle.
Increasing $\gbar s$ moves the cubic $v_{\infty}$ with the ensuing unstable fixed point $p_{f}$ down the sigmoid $w_{\infty}$ in the $(v-w)$-plane (\cref{fig:nullclines-coupled}).
When $\gbar s$ is large enough, the fixed point $p_{f}$ becomes stable, attracting all previously periodic trajectories.
There exists a unique value $\gbar s=g^{\star}$ when $p_{f}$ changes stability and the stable limit cycle vanishes.
Thus, when a constant inhibitory conductance is applied, $\gbar s<g^{\star}$ acts as a necessary condition for a cell to spike.
In contrast, when $\gbar s>g^{\star}$ inhibition is strong enough to prevent a cell from spiking~\citep{bose2011}.

% NOTE
% In our parameter range the bifurcations occur via subcritical Hopf (bose2011 p.47)

Now let us analyse the nullclines of the quiet cell when the two cells are coupled via synaptic inhibition with depression.
Let $\gbar s$ here denote the total synaptic conductance which acts on the quiet cell and is produced by the free cell, and let $p_{f}$ be the fixed point associated with the quiet cell.
At the start of the burst of the free cell we have $\gbar s > g^{\star}$ and $p_{f}$ is stable.
When the free cell spikes, $\gbar s$ peaks (\cref{eq:s-reset}), and the $v$-nullcline with the ensuing stable $p_f$ move down the $w$-nullcline.
Then in between spikes $\gbar s$ decays exponentially (\cref{eq:dot-s}) causing the $v$-nullcline and $p_{f}$ to move up the $w$-nullcline while attracting trajectories of the quiet cell.
Once depression causes the synaptic conductance to become small enough to satisfy $\gbar s<g^{\star}$ and the quiet cell is released, fixed point $p_{f}$ becomes unstable allowing the quiet cell to fire.
If the trajectory of the quiet cell remains sufficiently close to the stable $p_{f}$ at the time when it changes stability, then $\gbar s < g^{\star}$ acts as a release condition that is not only necessary, but also sufficient for firing of the quiet cell.
In this case the release of the quiet cell occurs precisely when
\begin{equation}
 ~\label{eq:release}
  \gbar s=g^{\star}
\end{equation}
is satisfied.

Whether the $(v,w)$-trajectory of the quiet cell can remain close enough to $p_{f}$ to make \cref{eq:release} sufficient for firing depends largely on the coupling strength $\gbar$ and the timescale disparity between membrane dynamics and synaptic dynamics~\citep{bose2011}.
It is straightforward to test our assumption of a release condition by numerically integrating the full system of ODEs and calculating the time interval between the first spike of the quiet cell and the time when $\gbar s$ first crosses $g^{\star}$.
We will call this time interval the ``release delay''.
If our assumption holds, we would expect an approximately zero release delay.
\Cref{fig:release-delay} shows the numerically computed graph of the release delay for varying $\gbar$.
The graph shows three distinct branches, next to each branch we also plot the timecourse of a corresponding sample solution of the total synaptic conductance $\gbar s$ of both cells.
For the rightmost branch where $\gbar>0.592$ $\si{mS/cm^{2}}$ the release delay is approximately zero.
Here the first spike of the quiet cell can be accurately predicted by the release condition in \cref{eq:release}.
The leftmost and middle branches for $\gbar<0.592$ $\si{mS/cm^{2}}$  show a release delay greater than zero, the quiet cell does not immediately fire when the release condition is first satisfied, and \cref{eq:release} does not accurately predict the release of the quiet cell.
The leftmost and middle branches contain $1-1$ and $2-2$ solutions respectively.
In both cases the coupling $\gbar$ is not sufficiently strong to allow trajectories of the quiet cell to be close enough to $p_{f}$ to guarantee spiking once $\gbar$ crosses $g^{\star}$.
Note that in the middle branch $\gbar s $ crosses $g^{\star}$ twice, and only after the second crossing does the quiet cell fire.
The following map construction relies on the assumption that the release condition in \cref{eq:release} can accurately predict the release time of the quiet cell.
Given our model parameters this is only possible for sufficiently large $\gbar>0.592 $ $\si{mS/cm^{2}}$.
For completeness, however, we will also consider values $\gbar < 0.592$ $\si{mS/cm^{2}}$ in the following analysis, bearing in mind that in this parameter range our map will not be accurate.

In summary: For $\gbar>0.592$ $\si{mS/cm^{2}}$ the release condition is sufficient to predict when the quiet cell is released.
Due to the symmetry of $n-n$ solutions the release occurs at exactly half the period of the full cycle, that is at $P/2$.
The release time therefore uniquely determines the type of $n-n$ solution.
Furthermore, computation of the release time does not depend on the membrane nor the synaptic dynamics of the quiet cell.
Instead, the solution of the synaptic variable $s$ of the free cell is sufficient to predict when $\gbar s=g^{*}$ is satisfied.
Finally, $s$ is solely determined by the evolution of the depression variable $d$ of the free cell.
Constructing a solution of $d$ during the free phase of either cell will therefore uniquely determine the solution of the full eight-dimensional network.
However, finding the solution $d$ requires us to know the initial value $d(0)$ at the start of a cycle at $t=0$.
In the next section we will construct a scalar return map that tracks these initial values $d(0)$ from cycle to cycle of stable $n-n$ solutions.

\subsection{Construction of the scalar Poincaré map}
In this section we construct the scalar Poincaré map $\Pi_{n}:d^{\star}\mapsto d^{\star}$.
Here the discrete variable $d^{\star}$ tracks the values of the continuous depression variable $d$ at the beginning of each $n-n$ burst.
The map $\Pi_{n}$ therefore describes the evolution of $d$, of either of the two cells, from the beginning of one cycle to the beginning of the next cycle.
To simplify the map construction we will assume that a free cell fires exactly $n$ times before it becomes quiet.
Later we will relax this assumption.
We will construct $\Pi_{n}$ by evolving $d$ first during the free phase and then during the quiet phase of the $n-n$ limit cycle.
First, let us give explicit definitions of the free and quiet phases.
A schematic illustration of both phases is given in \cref{fig:free-quiet1}.

Suppose that at $t=0$ cell 1 becomes free with some initial $d(0)$. 
Cell 1 then fires $n$ spikes at the uncoupled period $T$.
Let $s(t)$ and $d(t)$ be the corresponding solutions of the synaptic and depression variables of cell 1.
After $n$ spikes the total conductance $\gbar s(t)$ acting on the quiet cell 2 has decayed sufficiently to satisfy the release condition~\eqref{eq:release}, that is at some time $t=(n-1)T+\delt$ we have $\gbar s(t)=g^{\star}$.
Cell 2 is then released and prevents cell 1 from further spiking.
Here $\delt$ is the \emph{inter-burst-interval}, or the time between the last spike of cell 1 and the first spike of cell 2~\citep{bose2011}.
Once released, cell 2 also fires $n$ spikes until cell 1 becomes free once again at the cycle period.
Let $P_n$ denote the full cycle period of a $n-n$ solution:
\begin{equation}
 ~\label{eq:P}
  P_n = 2(n-1)T + 2\delt.
\end{equation}
We can now define the free and quiet phases of cell 1 explicitly. The free phase is the time interval between the first and last spikes of the burst, that is for time $0<t<(n-1)T$.
During the free phase of cell 1, the quiet cell 2 is inhibited sufficiently strong to prevent it from firing, hence $\gbar s > g^{\star}$.
The quiet phase of cell 1 is the remaining duration of the cycle when the cell is not firing, that is for $(n-1)T < t < 2(n-1)T + 2\delt$.

Note that only the quiet phase depends on $\delt$, and $\delt$ will play a central role in the construction of $\Pi_{n}$.
From \cref{eq:P} $\delt$ can be be computed as
\begin{equation}
 ~\label{eq:delta-t-P}
  \delt = \frac{1}{2}P_n - (n-1)T.
\end{equation}
\noindent
We can use \cref{eq:delta-t-P} and the numerically computed bifurcation diagram of the period for stable $n-n$ solutions in \cref{fig:bif-diagram}A to obtain the graph of $\delt$ as a function of $\gbar$ (\cref{fig:delta-t}).
Each continuous branch of $\delt$ is monotonically increasing and corresponds to a $n-n$ burst:
Stronger coupling $\gbar$ increases the total synaptic conductance $\gbar s$ that acts on the quiet cell, thus delaying its release.
It is easy to see that for any $n$-branch we have $\delt<T$:
Once $\delt$ crosses $T$, the free cell can ``squeeze in" an additional spike and the solutions bifurcate into a $(n+1)-(n+1)$ burst.

Distinguishing between the free and quiet phases of a cycle allows us to describe the dynamics of the depression variable $d$ explicitly for each phase.
As can be seen from \cref{fig:free-quiet1}C, during the free phase $d$ depresses at spike times and recovers in the $ISI$s.
In contrast, during the quiet phase $d$ only recovers and does not depress.
Given the initial $d^{\star}=d(0)$ at the beginning of the cycle and the number of spikes in the free phase $n$, we can now construct the burst map $\Pi_{n}$.
The map
\begin{equation}
  \Pi_{n}(d^{\star})=Q_{n}\big(F_{n}(d^{\star}\big))
\end{equation}
\noindent
is a composition of two maps.
Map
\begin{equation}
  F_{n}:d^{\star}\mapsto \delt
\end{equation}
models the evolution of $d$ in the free phase.
$F_{n}$ takes an initial value $d^{\star}$ and calculates the inter-burst-interval $\delt$.
Map
\begin{equation}
  Q_{n}:\delt \mapsto d^{\star}
\end{equation}
models the recovery of $d$ in the quiet phase.
Given some $\delt$ map $Q_n$ computes $\dstar$ at the start of the next cycle.

Our aim in the following analysis is to elucidate the properties of $\Pi_{n}$ and to understand the structure of its parameter space by exploring how the stable and unstable fixed points of $\Pi_{n}$ are created.
To that effect it will be useful to include not only positive, but also negative values of $d^{\star}$ to the domain of $\Pi_{n}$.
But it is important to add that values $d^{\star}<0$ are biologically impossible as the depression variable models a finite pool of neurotransmitters, and therefore must be positive.
Because $\Pi_{n}$ maps first from $d^{\star}$ to $\delt$, and then back to $d^{\star}$, we will also consider negative values of $\Delta t$, interpreting them as $n-n$ solutions with partially overlapping bursts.
As will become evident, $\delt<0$ is only a formal violation of the biological realism of the map $\Pi_{n}$, as numerically stable $n-n$ solutions of the full system of ODEs only exist for $\Delta t>0$.

We start the construction of $\Pi_n$ by first considering the free phase and building the map $F_n$.
At each spike time $t_{k}$, variable $d$ depresses with the reset $d(t_{k})\leftarrow \lambda d(t_{k}^{-})$ (\cref{eq:d-reset}).
For the duration of the subsequent $ISI$, the recovery of $d$ is described by the solution to \cref{eq:dot-d}:
\begin{equation}
  d(t) = 1 - (1-\lambda d_{k})e^{-t/\tau_{d}}~\label{eq:d-sol},
\end{equation}
where the variable $d_{k}=d(t^{-}_{k})$ is the level of depression at spike time $t_{k}$.
By substituting $t=T$ we can build a linear map that models the depression of $d$ from spike time $t_{k}$ to the subsequent spike time $t_{k+1}$ during the free phase:
\begin{equation}
  d_{k+1} = \lambda\rho d_{k} + (1-\rho),~\label{eq:map-d}
\end{equation}
where to keep the notation simple we let $\rho=\exp(-T/\tau_{d})$.
Since $0<\lambda, \rho<1$, the map~\eqref{eq:map-d} is increasing and contracting, with a fixed point at
\begin{equation}
 ~\label{eq:dsup}
  d_{s}=\frac{1-\rho}{1-\lambda\rho},
\end{equation}
where $0<d_{s}<1$.
The value $d_{s}$ is the maximum depression value that can be observed in the suppressed solution where the active cell fires at its uncoupled period $T$ (see \cref{fig:burst-sols}E).
Using the release condition in \cref{eq:release} allows us to derive the value of the the minimum coupling strength that will produce the full suppressed solution, denoted as $\gbar_{s}$.
Solving \cref{eq:dot-s} for $s(t)$ with $t=T$ and setting the initial value $s(0)=d_{s}$ gives
\begin{equation}
\gbar_{s} d_{s}e^{-T/\tau_{s}}=g^{\star}.
\end{equation}
By further substituting the definition of $d_{s}$ in~\eqref{eq:dsup} and rearranging, we can write $\gbar_{s}$ as a function of the depression strength $\lambda$:
\begin{equation}
  \label{eq:gs}
  \gbar_{s}(\lambda) = g^{\star}e^{T/\tau_{s}}\frac{1-\lambda \rho}{1-\rho}.
\end{equation}
Note that the above dependence of $\gbar_{s}$ on $\lambda$ is linear and monotonically decreasing.
Increasing $\lambda$ reduces the strength of the depression of the free cell.
This in turn allows the free cell to fully suppress the quiet cell at smaller values of $\gbar$.

Solving~\eqref{eq:map-d} gives us the linear map $\delta_{n}:d^{\star}\mapsto d_{n}$, that for some initial $d^{\star}$ computes the depression at the $n$th spike time, $d_{n}=d(t_{n}^{-})$:
\begin{equation}
 ~\label{eq:delta-map}
  \delta_{n}(d^{\star}) = (\lambda \rho)^{n-1} d^{\star} + (1-\rho)\sum_{i=0}^{n-2}(\lambda \rho)^{i}.
\end{equation}
Since $\lambda < 1$, function $\delta_{n}$ is a linearly increasing function of $d^{\star}$ with a fixed point at $d_{s}$ for all $n$.
Having identified $d$ after $n$ spikes, we can now use the release condition $\gbar s = g^{\star}$ (\cref{eq:release}) to find $\delt$.
At the last spike of the free phase at time $t_{n}=(n-1)T$ the synapse variable $s$ is reset according to \cref{eq:s-reset}:
\begin{equation}
  s(t_{n})\leftarrow \dn(\dstar).
\end{equation}
After the last spike $s$ decays exponentially according to the solution of \cref{eq:dot-s}, where setting the initial condition $s(0)=\dn(\dstar)$ yields:
\begin{equation}
~\label{eq:s-sol}
  s(\delt)=\dn(\dstar)e^{-\Delta t/\tau_{s}}.
\end{equation}
Substituting $s(\delt)$ into $s$ of the release condition (\cref{eq:release}) gives then
\begin{equation}
 ~\label{eq:release2}
  \gbar \dn(\dstar) e^{-\delt/\tau_{s}}=g^{\star}.
\end{equation}
Our assumption of the release condition guarantees that the quiet cell 2 spikes and becomes free when $\gbar s - g^{\star}$ crosses zero.
Solving \cref{eq:release2} for $\delt$ allows us to compute the inter-spike-interval as a function of $\dstar$, which defines our map $F_n$:
\begin{equation}
 ~\label{eq:Fn-map}
  F_{n}(d^{\star}):=\tau_{s}\ln{\left(\frac{\gbar }{g^{\star}} \delta_{n}(d^{\star})\right)}= \delt.
\end{equation}

\Cref{fig:FQ-map}A shows $F_n$ for various $n$, which is a strict monotonically increasing function of $d^{\star}$ as well as $\gbar$.
Larger values of $d^{\star}$ and $\gbar$, respectively, cause stronger inhibition of the quiet cell, and therefore prolong its release time and the associated $\delt$.
Map $F_{n}$ is defined on $d^{\star}>d_{a}$, where $d_{a}$ is a vertical asymptote found by solving $\delta_{n}(\dstar)=0$ in \cref{eq:delta-map} for $\dstar$, which yields
\begin{equation}
  d_{a}(n)=-\frac{(1-\rho)\sum_{i=0}^{n-2}(\lambda \rho)^{i}}{ (\lambda \rho)^{n-1} }\leq 0~\label{eq:da}.
\end{equation}

>>>>

\documentclass[../manuscript.tex]{subfiles}
\graphicspath{{../figures/}}

\begin{document}
% DEPRESSION
Short-term synaptic plasticity may have a role in burst activity in central pattern
generators (CPGs).  Short-term synaptic depression is commonly found in neuronal networks
involved in the generation of rhythmic movements, such as in the pyloric CPG of the spiny
lobster~\citep{manor1997, rabbah2007}, or in the lumbosacral cord of the chick
embryo~\citep{donovan1998}.  Synaptic depression modulates the strength of synapses in
response to changes to the presynaptic firing frequency.  At a high neuronal firing
frequency, depression weakens the strength of synapses and therefore reduces the magnitude
of the postsynaptic response.  At low firing frequency, it allows sufficient time for the
synapse to recover from depression between spikes, leading to a stronger postsynaptic
response.  In reciprocal networks, synaptic depression has been shown to act as a
``switch'', giving rise to a wide range of network dynamics such as synchronous and
multi-stable rhythms, as well as fine tuning the frequency of network
oscillations~\citep{nadim2000, nadim1999, bose2011}.

@@
